Hydras \& Co.

### 2022-02-09 doc/gaia-chapter.tex
re-redaction of chapter 10 (start)

\chapter{Gaia and the hydras (draft)}
\index{coq}{Plug-ins!Gaia}

@@

As an initial step in bridging \gaia and \Hydras, we were able to establish a
correspondence between both implementations of ordinals through rewriting
lemmas. For instance, we proved that multiplication of ordinals in Cantor
normal form in \Hydras refine \gaia's corresponding multiplication operation,
and then imported \gaia's proof of associativity of the multiplication
almost for free.

The initial draft bridge code mostly uses the SSReflect proof language
and idioms from the Mathematical Components library. We made this design
decision since we believe it is less challenging to reason about
Hydra-battles code using SSReflect and MathComp than to reason about
Gaia without SSReflect.

\section{Data refinement}

\emph{The following definitions and lemmas are in
\href{../theories/html/gaia_hydras.T1Bridge.html}%
{\texttt{gaia\_hydras.T1Bridge}}}.

Since both libraries \texttt{gaia} and \texttt{hydra-battles} have their own type \texttt{T1}, we will work with the following notations.

@@

@@

From now on, names like  \texttt{T1}, \texttt{omega}, etc. are  bound to \Hydras's meaning.

@@

\subsection{Porting theorems from \Hydras}

@@

First, we rename \Hydras's \texttt{canon} function into
\texttt{hcanon}. Then we define a new function \emph{via} the
translations \texttt{g2h} and \texttt{h2g}.

@@

The following lemmas are proved by rewriting from corresponding statements proved in \Hydras.
\emph{The author of these proof scripts is still a beginner in Ssreflect. Please forgive the clumsiness of the scripts below.}

@@

\section{Gaia's proof of well-foundedness}
The library \Hydras already contains two proofs
of well-foundedness of the strict order on ordinal terms in
Cantor normal form: a direct proof~\vref{sec:T1Wf} and a proof based on the recursive path ordering~\vref{remark:a3pat}, by
\'Evelyne Contejean. We found it interesting to comment Gaia's proof, which uses tools already present in~\cite{CantorContrib}.

@@

\emph{Notes: the abbreviation \texttt{LT}, and the \texttt{fold LT} and \texttt{unfold LT} tactics,are not in \texttt{gaia}, but  have been added to make the goal display more readable.
The infix notation \texttt{`` \_ < \_''}  is bound to the order \texttt{T1lt} in \gaia, while,  in \Hydras,  it is bound to its restriction \texttt{LT} to terms in normal form. }

>>>>

\chapter{Gaia and the hydras (experimental)}
\index{coq}{Plug-ins!Gaia}

@@

The bridge we aim to build and cross is a two-way bridge:

\begin{itemize}
\item From \gaia, we import the proof of associativity of multiplication of ordinals strictly less than $\epsilon_0$, which was missing from \Hydras.
  \item The other way, we started to import definitions and properties of canonical sequences and rapidly growing functions from \Hydras (see Chapter~\ref{chap:ketonen} and\ref{chap:alpha-large}~of this book).
\end{itemize}

Please note that by ``import'' we do not mean ``duplicate the proofs''.  Whenever possible, we proceed by rewriting steps over the statements and not the proofs.

This initial draft bridge code mostly uses the SSReflect proof language
and idioms from the Mathematical Components library. We made this design
decision since we believe it is less challenging to reason about
Hydra-battles code using SSReflect and MathComp than to reason about
Gaia without SSReflect.

\section{Library structure}
The \gaiaHydras library is designed for use  by users of
\gaia and/or  \HydrasLib. Whenever possible, we managed to respect the notations and naming conventions of these libraries.

In case of name clash, we often added a one-letter prefix to names : \texttt{g} or \texttt{G} for \gaia, \texttt{h} or \texttt{H}
  for \HydrasLib.

For instance, both \gaia and \HydrasLib have a type called \texttt{T1} (data type for Cantor normal forms).
\begin{itemize}
\item In
  \href{../theories/html/gaia_hydras.T1Bridge_use.html}%
  {\texttt{gaia\_hydras.T1Bridge\_use}}, we adopt \HydrasLib's point of view, and the name \texttt{T1} is by default bound to
\HydrasLib' data structures.

\inputsnippets{T1Bridge_use/T1BridgeUse,
  T1Bridge_use/LocateT1}

Associativity of multiplication over \HydrasLib' type \texttt{T1},
although imported from \gaia, is expressed with \HydrasLib' notations.

\inputsnippets{T1Bridge_use/T1BridgeUseb}

\item The other way, the module \href{../theories/html/gaia_hydras.GCanon.html}%
  {\texttt{gaia\_hydras.GCanon}} is an adaptation to \gaia of
  \href{../theories/html/hydras.Epsilon0/Canon.html}%
  {\texttt{hydras.Epsilon0/Canon}}.
  In this module, the name \texttt{T1} is bound by default to \gaia's meaning. \HydrasLib' name \texttt{T1} is renamed into \texttt{hT1}.

  The following lemmas are expressed in \gaia's vocabulary.
  \inputsnippets{GCanon/gcanonLimitMono,
    GCanon/gcanonLimitOf}

\end{itemize}

The \gaiaHydras entry point is the file
\href{../theories/html/gaia_hydras.T1Bridge.html}%
{\texttt{theories/gaia.T1Bridge.v}}, which requires modules both from \HydrasLib and \gaia. By default, we give priority to \gaia's notation and vocabulary.

\inputsnippets{T1Bridge/Requirements}

@@

\subsection{Data refinement}

@@

From now on, names like  \texttt{T1}, \texttt{omega}, etc. are  bound to \HydrasLib' meaning.

@@

\subsection{Porting theorems from \HydrasLib}

@@

First, we rename \HydrasLib' \texttt{canon} function into
\texttt{hcanon}. Then we define a new function \emph{via} the
translations \texttt{g2h} and \texttt{h2g}.

@@

The following lemmas are proved by rewriting from corresponding statements proved in \HydrasLib.
\emph{The author of these proof scripts is still a beginner in Ssreflect. Please forgive the clumsiness of the scripts below.}

@@

\section{Gaia's proof of well-foundedness}
The library \HydrasLib already contains two proofs
of well-foundedness of the strict order on ordinal terms in
Cantor normal form: a direct proof~\vref{sec:T1Wf} and a proof based on the recursive path ordering~\vref{remark:a3pat}, by
\'Evelyne Contejean. We found it interesting to comment Gaia's proof, which uses tools already present in~\cite{CantorContrib}.

@@

\emph{Notes: the abbreviation \texttt{LT}, and the \texttt{fold LT} and \texttt{unfold LT} tactics,are not in \texttt{gaia}, but  have been added to make the goal display more readable.
The infix notation \texttt{`` \_ < \_''}  is bound to the order \texttt{T1lt} in \gaia, while,  in \HydrasLib,  it is bound to its restriction \texttt{LT} to terms in normal form. }
